### 2014-03-29 LaTeX/RhodiumEmp.tex
more changes in empirics

To measure economic performance we use annual percent change in GDP, which we collect from the World Development Indicators at the World Bank. We choose to focus on year over change in GDP because we expect conflicts proximate to major cities
to lead to reductions in growth not changes in levels.\footnote{Alternative measures of economic performance would have included changes in FDI inflows or changes in GDP per capita, in using either of these measures are results remain relatively the same.} Unlike much of the extant literature in constructing our dependent variable we focus on year over change instead of a ten year average. We believe this approach to be superior because it allows us to estimate the direct effect of the conflict on economic performance in that year. A ten-year average approach might hide important variation in our dependent variable that can be explained by the proximity of conflict to economic centers.

@@

The key theory that we advance here is that the effects of conflict on economic performance are conditional on the proximity of those conflicts to major city centers. Thus a key part of the research here was determining major city centers by year. The only data source we found with enough geographical and temporal coverage on major cities was The World Alamanac. To gather data from these textbooks we hired several undergraduate students to code the major cities by country for each year of the Almanac from 1989 to 2008. Major cities are classified as those that have relatively large shares of a country's population and/or are economic centers. On average, for each country at least three major cities are identified in each year from 1989 to 2008. In figure \ref{fig:CityConfMap}, we show the geographic distribution of conflicts and cities.

\begin{figure}[ht]
	\centering
	\includegraphics[width=.9\textwidth]{CityConfMap-crop}
	\caption{This map illustrates the geographic distribution of all conflicts, according to the PRIO Conflict Site Dataset, and major cities listed in The World Almanac from 1989 to 2008. Countries for which no armed conflicts are recorded are shaded in grey.}
	\label{fig:CityConfMap}
\end{figure}

We restrict our sample to only those countries that are listed as having conflicts in the PRIO dataset. As a result, our analysis focuses on exploring variations in economic growth in countries undergoing conflict by the proximity of those conflicts to major cities. To determine proximity, we calculate the distance in kilometers between the centroid of each of the conflicts with the centroids of city locations. Thus for each year we determine the distance between every conflict and every major city within the state. For example, in figure \ref{fig:columbiaMap} we show the location of armed conflicts in Colombia, circles filled in shades of red, and the major cities therein, blue diamonds.\footnote{One armed conflict involving rebel actors and the Colombian government occurred outside of Colombia's borders.} For Colombia each of these points represents the same conflict but a different year, with earlier years shaded in light red and later years shaded in dark red. Thus for each year we calculated the distance between the centroid conflict location provided by PRIO and every major city identified in blue.

\begin{figure}[ht]
	\centering
	\includegraphics[width=.7\textwidth]{colombiaMap-crop}
	\caption{This map illustrates the geographic distribution of all conflicts, according to the PRIO Conflict Site Dataset, and major cities listed in The World Almanac from 1989 to 2008. Countries for which no armed conflicts are recorded are shaded in grey.}
	\label{fig:columbiaMap}
\end{figure}

A key problem we face in conducting this analysis is reconciling the unit of analysis for our key independent and dependent variables. Our dependent variable is at the country-year and our key independent variable, distance between conflicts and major cities, are at the conflict-country-year level. To deal with this, we aggregate up to the country-year level by calculating the minimum logged distance any conflict is from a major city. Thus if a country faced, for example, four conflicts in one year, the datapoint that we would aggregate up to the country-year level would be for the conflict that was closest to a major city. This obviously is somewhat problematic as we end up discarding the information from the other conflicts. At the same time, we argue that this choice of aggregation is what conforms closest to our theoretical claims about economic activity only being severely disrupted in cases of conflicts being proximate to major cities. From the PRIO dataset we also bring in additional information about that conflict, specifically:

>>>>

To measure economic performance we use annual percent change in GDP, which we collect from the World Development Indicators at the World Bank. We choose to focus on year over change in GDP because we expect conflicts proximate to major cities to lead to reductions in growth.\footnote{Alternative measures of economic performance would have included changes in FDI inflows or changes in GDP per capita, in using either of these measures are results remain relatively the same.} Unlike much of the extant literature in constructing our dependent variable we focus on year over change instead of a ten year average. We believe this approach to be superior because it allows us to estimate the direct effect of the conflict on economic performance in that year. A ten-year average approach might hide important variation in our dependent variable that can be explained by the proximity of conflict to economic centers.

@@

The key theory that we advance here is that the effects of conflict on economic performance are conditional on the proximity of those conflicts to major city centers. Thus a crucial part of our work here involved determining major city centers by year. The only data source we found with enough geographical and temporal coverage on major cities was The World Alamanac. To gather data from these textbooks we hired several undergraduate students to identify the major cities listed for each country within a yearly edition of the Almanac from 1989 to 2008. Major cities are classified as those that have relatively large shares of a country's population and/or are economic centers. Typically, the Alamanacs listed at least three major cities for each country and year from 1989 to 2008. After we had gathered a yearly list of major cities by country, we determined the centroid location of each city using Google's Geocoding API.

In figure \ref{fig:CityConfMap}, we show the geographic distribution of conflicts and cities. The centroid locations of conflicts are shown by red dots. Not surprisingly, we can see that in many cases conflict locations are clustered within specific parts of a country. In most cases, this clustering is indicative of the same conflict moving within the geographic boundaries of a country over time. The blue diamonds in \ref{fig:CityConfMap} denote the locations of major cities as identified in The World Alamanacs from 1989 to 2008. Countries shaded in grey are those for which no armed conflict took place in this period according to the PRIO dataset.

\begin{sidewaysfigure}[ht]
	\centering
	\includegraphics[width=.9\textwidth]{CityConfMap-crop}
	\caption{This map illustrates the geographic distribution of all conflicts, according to the PRIO Conflict Site Dataset, and major cities listed in The World Almanac from 1989 to 2008. Countries for which no armed conflicts are recorded are shaded in grey.}
	\label{fig:CityConfMap}
\end{sidewaysfigure}
\FloatBarrier

Our analysis focuses on exploring variations in economic growth in countries undergoing conflict by the proximity of those conflicts to major cities, thus we restrict our sample to only those countries that are listed as having conflicts in the PRIO dataset. To determine proximity, we calculate the distance in kilometers between the centroid of each of the conflicts with the centroids of city locations. Meaning that for each year we determine the distance between every conflict and every major city within a state. For example, in figure \ref{fig:columbiaMap} we show the location of armed conflicts in Colombia, dots colored in shades of red, and the major cities therein, blue diamonds.\footnote{One armed conflict involving rebel actors and the Colombian government occurred outside of Colombia's borders.} For Colombia each of the red dots represents the same conflict but a different year, with earlier years shaded in light red and later years shaded in dark red. Thus for every year we made six distance calculations -- one for every city.

\begin{figure}[ht]
	\centering
	\includegraphics[width=.5\textwidth]{colombiaMap-crop}
	\caption{This map illustrates the geographic distribution of all conflicts, according to the PRIO Conflict Site Dataset, and major cities listed in The World Almanac from 1989 to 2008. Countries for which no armed conflicts are recorded are shaded in grey.}
	\label{fig:columbiaMap}
\end{figure}

A key problem we face in conducting this analysis is reconciling the unit of analysis for our key independent and dependent variables. Our dependent variable is at the country-year and our key independent variable, distance between conflicts and major cities, is at the conflict-country-year level. To deal with this, we aggregate up to the country-year level by calculating the minimum logged distance any conflict is from a major city. Thus if a country faced, for example, four conflicts in one year, the datapoint that we would aggregate up to the country-year level would be for the conflict that was closest to a major city. This obviously is somewhat problematic as we end up discarding the information from the other conflicts. At the same time, we argue that this choice of aggregation is what conforms closest to our theoretical claims about economic activity only being severely disrupted in cases of conflicts being proximate to major cities. From the PRIO dataset we also bring in additional information about that conflict, specifically:
